---
title: "Gym Attendance Prediction"
author: "Olamide Adu"
date: "2023-08-07"
output:
   html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    dev: svg
    theme: simplex
    highlight: zenburn
    code_folding: show
    keep_md: true
    df_print: paged
---

# Goalzone Fitness Class: Predict the Attendance of Members

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Company Description

GoalZone is a fitness club chain providing fitness classes in Canada. Finally, the fitness classes schedule is back to normal after the COVID-19 restrictions are lifted. However, they have received many complaints from the customers about having a hard time booking a fitness class.

**The Problem**: Some classes are always fully booked, and fully booked classes often have a low attendance rate.

Goalzone wants to increase the number of spaces available for classes. They want to do this by predicting whether the member will attend the class or not. If they can predict a member will not attend the class, they can make another space available

## Customer Question

-   The operation manager has asked to answer the following: - Can you predict the member that will attend a class or not?

## Dataset

The dataset contains the attendance information for the class scheduled this year so far. The data comes from DataCamp as a part of their Data Science Certificate Project.

| Column Name      | Criteria                                                                                                                                                            |
|--------------------------------|----------------------------------------|
| booking_id       | Nominal. The unique identifier of the booking. Missing values are not possible due to the database structure.                                                       |
| months_as_member | Discrete. The number of months as this fitness club member, minimum 1 month. Replace missing values with the overall average month.                                 |
| weight           | Continuous. The member's weight in kg, rounded to 2 decimal places. The minimum possible value is 40.00 kg. Replace missing values with the overall average weight. |
| days_before      | Discrete. The number of days before the class the member registered, minimum 1 day. Replace missing values with 0.                                                  |
| day_of_week      | Nominal. The day of the week of the class. One of "Mon", "Tue", "Wed", "Thu", "Fri", "Sat" or "Sun". Replace missing values with "unknown".                         |
| time             | Ordinal. The time of day of the class. Either "AM" or "PM". Replace missing values with "unknown".                                                                  |
| category         | Nominal. The category of the fitness class. One of "Yoga", "Aqua", "Strength", "HIIT", or "Cycling". Replace missing values with "unknown".                         |
| attended         | Nominal. Whether the member attended the class (1) or not (0). Missing values should be removed.                                                                    |

Load libraries, import data and preview data

```{r, libraries, results='hide', message = FALSE}
library(tidyverse)
library(ggthemes)
library(rsample)
library(pROC)
library(rpart)
library(rpart.plot)

# Import data
goalzone_fc <- read_csv("https://raw.githubusercontent.com/xrander/fitness_class_project/master/fitness_class_2212.csv")

```

```{r}
goalzone_fc #view document
```

## Data Validation
```{r, warning = FALSE, message=FALSE}
## Check the structure of the data to see if column follow the data description
str(goalzone_fc)
```
